% Created 2019-01-28 Mon 22:05
% Intended LaTeX compiler: pdflatex
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{minted}
\usepackage{tikzsymbols}
\usetheme{Madrid}
\author{Jakub Zárybnický}
\date{2019-01-29}
\title{A Haskell Platform for Creating Progressive Web Applications}
\subtitle{A midterm progress report}
\title[PWAs in Haskell]{A Haskell Platform for Creating Progressive Web Applications}
\hypersetup{
 pdfauthor={Jakub Zárybnický},
 pdftitle={A Haskell Platform for Creating Progressive Web Applications},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.1 (Org mode 9.1.9)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{What, why?}
\label{sec:org214f972}
\begin{frame}[label={sec:org6214eda}]{Frontend Haskell?}
\begin{itemize}
\item What's Haskell?  \pause
\begin{itemize}
\item \emph{"Purely functional programming language with type inference and lazy
evaluation"}  \pause
\item \emph{"If it compiles, it works."}   \pause
\end{itemize}
\item Why Haskell in the browser?     \pause
\begin{itemize}
\item JavaScript doesn't have a monopoly on the frontend
\begin{itemize}
\item Compile-to-JS (see TypeScript, see Elm)
\end{itemize}
\item Code sharing between server and client (business logic, entities)
\end{itemize}
\end{itemize}
\pause
\begin{itemize}
\item Many companies use it, several contribute to OSS:
\begin{itemize}
\item Obsidian - Reflex, Obelisk, Rhyolite
\item Tweag - Asterius, Inline-js
\item QFPL - Reflex-workshop, many UI components
\item IOHK (Cardano) - iohk-ops
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org130a101}]{PWA?}
\begin{itemize}
\item New buzzword from Google
\item Recent web development trend
\item An \emph{almost} native app
\begin{itemize}
\item Load a website
\item Click "Add to home screen"
\item Use like a native application
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org1ce6ebf}]{PWA features}
\begin{itemize}
\item \emph{Progressive Web App Checklist} -- Google
\end{itemize}
\pause
\begin{itemize}
\item "Basic PWA":
\begin{itemize}
\item Responsive and fast ("Native experience")
\item Available offline
\item Has "Add to home screen"
\item Push notifications
\item Device APIs
\end{itemize}
\end{itemize}
\pause
\begin{itemize}
\item "Exemplary PWA":
\begin{itemize}
\item Pre-rendered
\item Social share, metadata
\item Accessible
\item Full-screen mode
\item \ldots{}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:orgb5779f4}]{PWA technologies}
\begin{itemize}
\item Technologies:
\begin{itemize}
\item Service Workers
\item Web App Manifest
\item History API
\item Web Share API
\item Network Information API
\item Credential Management API
\end{itemize}
\end{itemize}
\end{frame}

\section{What have I done?}
\label{sec:org9677829}
\begin{frame}[label={sec:orgbca29ce}]{What have I been doing?}
\begin{itemize}
\item Research:
\begin{itemize}
\item Webapp frameworks across languages
\begin{itemize}
\item PWA developers' expectations
\end{itemize}
\item Available Haskell libraries and tooling    \pause
\item \(\Rightarrow\) missing tools and libraries
\item \(\Rightarrow\) only some in scope of my thesis
\end{itemize}
\end{itemize}

\pause

\begin{itemize}
\item Prototypes:
\begin{itemize}
\item A full-stack application
\item An offline-capable application
\item A frontend debugger toolbar
\item CI/CD for Haskell
\item \ldots{}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org560062b}]{What haven't I been doing?}
\pause

\begin{itemize}
\item Regular writing
\begin{itemize}
\item Blog
\item Progress reports   \pause
\begin{itemize}
\item Oops\ldots{}
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\section{What's next?}
\label{sec:org7a59691}
\begin{frame}[label={sec:org1d88469}]{What's next?}
\begin{itemize}
\item Must-haves:
\begin{itemize}
\item Browser routing (client library)
\item Service Workers (build tool and/or client library)
\item Push notifications (server \& client library)
\end{itemize}
\end{itemize}
\pause
\begin{itemize}
\item Nice-to-haves:
\begin{itemize}
\item Pre-rendering (server library)
\item Data sync (server \& client library)
\end{itemize}
\end{itemize}
\pause
\begin{itemize}
\item Four months should be enough, right? \Smiley
\end{itemize}
\end{frame}

\begin{frame}[label={sec:orge53a0c0}]{Wrapping up}
\begin{itemize}
\item Thanks for listening!
\end{itemize}

\begin{itemize}
\item \url{https://github.com/zarybnicky/thesis}
\end{itemize}
\end{frame}
\end{document}